---
title: "Fit GAM model of spore counts"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float:
      collapsed: no
      smooth_scroll: yes
    code_folding: hide
    toc_depth: 2
    css: style.css
editor_options: 
  chunk_output_type: console
---

```{r markdown-settings, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(
  out.width = '60%',
  fig.width = 7,
  fig.height = 7,
  message = FALSE,
  warning = FALSE
)
```

## Load libraries

```{r session settings}
library("tidyverse")
library("ggpubr")
library("gratia")
library("here")
library("mgcv")
theme_set(theme_pubclean())
```

## Spore proportion chi-squared analysis

Check if the spore type, number of nuclei, affects the dispersal.

```{r x-sq}
set.seed(3)
read_csv(here("data/spore_type.csv"),
         show_col_types = FALSE) %>%
  group_by(meter, spore_nuclei) %>%
  summarise(n = sum(count), .groups = 'drop') %>%
  spread(spore_nuclei, n) %>%
  column_to_rownames('meter') %>%
  chisq.test(simulate.p.value = TRUE) 
```

## Spore dispersal GAM Model

### Load data

```{r load-data}
load(here("data/mod_dat_juan.Rdata"))
load(here("data/newd.Rdata"))
```

### Inspect the mod_dat object

```{r inspect-mod-dat}
mod_dat %>% #str
  distinct(field, trap_coord, .keep_all = TRUE) %>%
  select(trap_coord,
         trap_degrees,
         degree_dif,
         wind_degrees,
         wind_speed,
         distance_m) %>%
  print(n = Inf)
```

```{r}
mod_dat %>% 
  ggplot()+
  aes(x=spore_cm2)+
  geom_histogram(fill="steelblue")+
  theme_bw()

```

Based on the use of AIC to evaluate model fit to the data, the following model best fit the data with smoothed terms for x and y, the Cartesian coordinates of the spore traps placed around the harvest fields, `x` and `y`; the smoothed time slice variables, `time_slice`; the wind direction in relation to the traps' placement, `degree_dif`; the traps' distance from the harvest field, `distance_m`; and the field and traps themselves as a random effects, `field` and `xy`.

```{r fit-m1, eval=FALSE}
mod_dat$xy <- as.factor(paste(mod_dat$x, mod_dat$y))
# mod_dat %>% distinct(x,y) %>% data.frame()
# mod_dat %>% rio::export("data/data_mod.xlsx")

m1 <- gam(
  spore_cm2 ~ s(y, x, k = 20) +
    # s(wind_speed, k = 3) +
    s(time_slice, k = 3) +
    s(degree_dif, k = 30) +
    s(distance_m, k = 4) +
    s(field, xy, bs = "re"),
  data = mod_dat,
  select = TRUE,
  method = "REML",
  family = "tw"
)
```

```{r save-m1, echo=FALSE, eval=FALSE}
save(m1, file = here("data/m1_juan.Rdata"))
```

```{r read-m1, echo=FALSE}
load(here("data/m1_juan.Rdata"))
```

### Inspect Model m1

```{r inspect-m1}
m1
```

#### Check model

```{r gam-check-m1}
gam.check(m1)
summary(m1)
```

The model fully converged after 15 iterations.
All of the predictive factors included were significant at _P_ > 0.05 and the smoothed predictors indicate a good fit with residuals being randomly distributed and enough basis functions for them to be properly fit.
The Q-Q plot generally falls along the line indicating a sufficient fit.
There is a bit of a pattern in the residual vs linear predictors but the histogram of residuals is a nice bell-shape.
The response versus fitted values clusters nicely around the 1:1 line.
In all, this model looks to explain the spore dispersal data very well.

#### Check for autocorrelation

Because the data represent spatially and temporally correlated data, it's best to check for any autocorrelation in the residuals.

```{r gam-autocorrelation}
acf(residuals(m1))
pacf(residuals(m1))
```

This looks OK.
There is no major pattern discernible in the ACF or PACF plots that show cause for concern.

### Summarise and visualise model m1

```{r appraise-m1,  out.width = '100%', fig.height = 6}
draw(m1)
```

```{r save-m1-summary, eval=FALSE, echo=FALSE}
ggsave(
  last_plot(),
  file = "plots_manuscript/model_summary.png",
  width = 8,
  height = 5,
  units = "cm",
  dpi = 300,
  scale = 3
)
ggsave(
  last_plot(),
  file = "plots_manuscript/model_summary.eps",
  device = cairo_ps,
  fallback_resolution = 600,
  width = 8,
  height = 5
)
```

### Predict spore dispersal and visualise spore dispersal predictions

```{r predict, eval=FALSE}
dist <- seq(100, 400, by = 100)

newd <- expand_grid(
  time_slice = c(90, 180, 270),
  degree_dif = seq(0, 360, by = 45),
  distance_m = coords_df$distance_m,
  wind_speed = seq(0, 20, by = 10),
  x = coords_df$x,
  y = coords_df$y,
  field = 1)
  
newd$pred_spore_cm2 <- predict(m1, newd)
```

Visualise the spore dispersal predictions for traps at 0m, 100m, 200m, 300m, 400m around the harvest field for sampling times at 90, 180 and 270 minutes after harvest started.

```{r graph-predictions}
predictions <-
  newd %>%
  mutate(pred_spore_cm2 = replace(pred_spore_cm2, pred_spore_cm2 < 0, 0)) %>%
  ggplot(aes(x = as.factor(distance_m), y = pred_spore_cm2)) +
  geom_violin() +
  labs(x = "Distance from harvesting field (m)",
       y = "Predicted Spores / cm²")

predictions
```

```{r save-predictions, eval=FALSE, echo=FALSE}
save(newd, file = "data/newd.Rdata"))
```

```{r load-newd, echo=FALSE}
load(here("data/newd.Rdata"))
```

```{r save-prediction-graph, eval=FALSE}
ggsave(
  last_plot(),
  file = "plots_manuscript/spore_prediction.png",
  w = 5,
  h = 2,
  scale = 1.2
)
```
