---
title: "Visualise data"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float:
      collapsed: no
      smooth_scroll: yes
    code_folding: hide
    toc_depth: 2
    css: style.css
editor_options: 
  chunk_output_type: console
---

```{r markdown-settings, echo=FALSE}
knitr::opts_chunk$set(
  out.width = '60%',
  fig.width = 7,
  fig.height = 7,
  message = FALSE,
  warning = FALSE
)
```

## Load libraries

```{r session settings}
library("knitr")
library("tidyverse")
library("ggpubr")
library("here")
library("scales")
```


```{r plots settings}
theme_set(theme_pubclean())

central_top_legend <- theme(
    legend.position = "top",
    legend.justification = "center",
    legend.direction = "horizontal",
    legend.background = element_blank(),
    legend.key = element_blank()
  )
```

## Load data

```{r load-data}
load(here("data/dat.Rdata"))
load(here("data/mod_dat.Rdata"))
sdp_dat <- read_csv(here("data/sdp_data.csv"),
                    show_col_types = FALSE) %>% 
                    arrange(field)
```

## Plot severely damaged pod indices

Visualize peanut smut incidence in the fields.

```{r sdp-incidence}
ggplot(data = sdp_dat) +
  aes(y = reorder(field, SDP), x = SDP) +
  geom_boxplot() +
  ylab("Field")
```

```{r sdp-incidence-mean_se}
ggline(sdp_dat,
       x = "field",
       y = "inc",
       add = c("mean_se", "jitter")) +
  xlab("Field") +
  ylab("SDP Incidence")
```

## Plot spore dispersal density

### Plot spore count densities

Create a density plot of the observed spore dispersal values.

```{r spore-kernel-density}
ggplot(mod_dat, aes(y = n_spore)) +
  geom_density()
```

Create a scatter plot and use `stat_smooth()` to fit a line.

```{r spore-dot-density-by-distance}
spore_dot_density_by_distance <- mod_dat %>%
  ggplot(aes(x = distance_m, y = n_spore)) +
  geom_point() +
  scale_fill_viridis_c() +
  geom_smooth(
    col = "grey50",
    method = "gam",
    formula = y ~ s(x, bs = "cs", k = 3)
  ) +
  labs(y = "Spores (n)",
  x = "Distance (m)") +
  facet_grid(. ~ time_slice)
spore_dot_density_by_distance
```

Create box plots showing spore distribution at the four distances.

```{r spore-violin-density-by-distance}
ggplot(mod_dat,
       aes(
         x = as.factor(distance_m),
         y = n_spore,
       )) +
  geom_boxplot() +
  scale_fill_viridis_d() +
  scale_colour_viridis_d() +
  labs(y = "Spores (n)",
       x = "Distance (m)")
```

```{r spore-violin-density-by-distance-by-field}
ggplot(mod_dat,
       aes(
         x = as.factor(distance_m),
         y = n_spore,
         colour = field,
         fill = field
       )) +
  geom_boxplot() +
  scale_fill_viridis_d() +
  scale_colour_viridis_d() +
  labs(y = "Spores (n)",
       x = "Distance (m)") +
  facet_grid(. ~ field)
```

```{r save-spores-density-violin, eval=TRUE, echo=FALSE}
ggsave(
  last_plot(),
  file = "plots_manuscript/spores_density_violin.png",
  width = 8,
  height = 5,
  units = "cm",
  dpi = 300,
  scale = 3
)
ggsave(
  last_plot(),
  file = "plots_manuscript/spores_density_violin.eps",
  device = cairo_ps,
  fallback_resolution = 600,
  width = 8,
  height = 5
)
```

Create a point plot showing spore density distribution over distance for each field.

```{r spore-distance-dispersal}
ggplot(data = mod_dat,
       aes(x = distance_m,
           y = n_spore,
           size = n_spore,
           colour = field)) +
  geom_point(alpha = 0.75) +
  scale_size(
    range = c(0.25, 6),
    name = "Smut Spores (n)",
    breaks = seq(10, 300, by = 60)
  ) +
  scale_colour_viridis_d() +
  labs(y = "Spores (n)",
  x = "Distance (m)") +
  central_top_legend
```


```{r save-spores-density-points, eval=TRUE, echo=FALSE}
ggsave(
  last_plot(),
  file = "plots_manuscript/spores_density_points.png",
  width = 8,
  height = 5,
  units = "cm",
  dpi = 300,
  scale = 3
)
ggsave(
  last_plot(),
  file = "plots_manuscript/spores_density_points.eps",
  device = cairo_ps,
  fallback_resolution = 600,
  width = 8,
  height = 5
)
```

## Filter and plot by downwind only


```{r downwind}
mod_dat$degree_com <-
  ifelse(
    mod_dat$degree_dif < -180, -360 + mod_dat$trap_degrees + mod_dat$wind_degrees,
    ifelse(
      mod_dat$degree_dif < 180,
      mod_dat$degree_dif,
      360 - mod_dat$trap_degrees + mod_dat$wind_degrees
    )
  )

mod_dat$degree_com <- abs(mod_dat$degree_com)

all_data <- mod_dat %>% 
  mutate_at(vars(time_slice), as.factor)

all_data$degree_group <-
  cut(
    all_data$degree_com,
    breaks = c(0, 45, 135, 180),
    labels = c("upwind", "cross wind", "downwind"),
    include.lowest = TRUE
  )

all_data2 <- all_data %>% filter(degree_group == "downwind")

ggplot(all_data2,
       aes(x = as.factor(distance_m), y = n_spore,
           fill = time_slice)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.60)) +
  scale_fill_viridis_d(name = "Time slice (min)") +
  labs(y = "Spores (n)",
       x = "Distance (m)") +
  labs(subtitle = "Smut spores trapped downwind") +
  facet_wrap(. ~ field,
              ncol = 3) +
  central_top_legend

ggsave(
  last_plot(),
  file = "plots_manuscript/spores_downwind.png",
  width = 8,
  height = 5,
  units = "cm",
  dpi = 300,
  scale = 3
)
ggsave(
  last_plot(),
  file = "plots_manuscript/spores_downwind.eps",
  device = cairo_ps,
  fallback_resolution = 600,
  width = 8,
  height = 5
)
```

#### Check windspeed effect

Create a hexbin plot to display the number of spores by wind speed and fit a smoothed line to these variables.

```{r spore-windspeed}
ggplot(data = mod_dat, aes(x = wind_speed, y = n_spore)) +
  scale_fill_viridis_c() +
  geom_hex() +
  geom_smooth(
    col = "grey50",
    method = "gam",
    formula = y ~ s(x, bs = "cs")
  ) +
  labs(y = expression(Spores / cm ^ {
    2
  }),
  x = "Wind Speed (m/s)") +
  central_top_legend +
  guides(fill = guide_legend(title.position = "top",
                             title.hjust = 0.5))
```

#### Polar coordinates

Create polar coordinate plots that display spore density in each cardinal direction.
Note that the 270 minute traps were deployed from time 0 to 270 minutes, 180 from 0 to 180, and 90 from 0 to 90.
So, here the figure shows an additive effect of the number of spores that were captured over time.
These are not discrete time_slices shown here, they do overlap.

```{r plot-spores,  out.width = '100%'}
ggplot(data = mod_dat) +
  aes(
    x = trap_degrees,
    y = distance_m,
    colour = n_spore,
    size = n_spore
  ) +
  facet_grid(time_slice ~ field) +
  coord_polar(theta = "x",
              start = 0,
              direction = 1) +
  geom_count(alpha = 0.55) +
  scale_colour_viridis_c(
    direction = -1,
    name = "Smut Spores (n)",
    guide = "legend",
    breaks = seq(10, 1410, by = 200),
    begin = 0,
    end = 0.8
  ) +
  scale_size(
    range = c(0.25, 6),
    name = "Smut Spores (n)",
    breaks = seq(10, 1410, by = 200),
  ) +
  scale_x_continuous(
    breaks = c(0, 90, 180, 270),
    expand = c(0, 0),
    limits = c(0, 360),
    labels = c("N", "E", "S", "W"),
    sec.axis = sec_axis(
      ~ . ,
      name = "Field",
      breaks = NULL,
      labels = NULL
    )
  ) +
  scale_y_continuous(
    breaks = c(0, 100, 200, 300, 400),
    limits = c(0, 400),
    sec.axis = sec_axis(
      ~ . ,
      name = "Minutes since harvest beginning",
      breaks = NULL,
      labels = NULL
    )
  ) +
  ylab("Distance (m)") +
  xlab("") +
  central_top_legend +
  guides(color = guide_legend(title.position = "top",
                              title.hjust = 0.5))
```

```{r save-spores-cardinal, eval=TRUE, echo=FALSE}
ggsave(
  last_plot(),
  file = "plots_manuscript/spores_cardinal.png",
  width = 8,
  height = 5,
  units = "cm",
  dpi = 300,
  scale = 3
)
ggsave(
  last_plot(),
  file = "plots_manuscript/spores_cardinal.eps",
  device = cairo_ps,
  fallback_resolution = 600,
  width = 8,
  height = 5
)
```

## Plot wind data

### Wind roses

<<<<<<< Updated upstream
Using the `wind_rose()` from _clifro_, create windrose plots to visualise windspeeds and directions during the harvest and spread events.
=======
Using a modified `wind_rose()` from {clifro}, create windrose plots to visualise wind speeds and directions during the harvest and spread events.
>>>>>>> Stashed changes

```{r wind-rose, out.width = '100%'}

<<<<<<< Updated upstream
# from SO: https://stackoverflow.com/questions/17266780/wind-rose-with-ggplot-r

library(scales)

facet = mod_dat$field
 calm_wind = 0
                    variable_wind = 990
                    n_col = 1
                    facet_row = mod_dat$time_slice

 optimal_n_dir = seq(1, 45, 2) * 4
  if (is.na(match(n_directions, optimal_n_dir))){
    n_directions = optimal_n_dir[which.min(abs(n_directions - optimal_n_dir))]
    message("using the closest optimal number of wind directions (",
            n_directions, ")")

  ## Create factor variable for wind direction intervals
  dir_bin_width = 360 / n_directions
  dir_bin_cuts = seq(dir_bin_width / 2, 360 - dir_bin_width / 2, dir_bin_width)
  dir_intervals = findInterval(c(direction, dir_bin_cuts), dir_bin_cuts)
  dir_intervals[dir_intervals == n_directions] = 0
  factor_labs = paste(c(tail(dir_bin_cuts, 1), head(dir_bin_cuts, -1)),
                      dir_bin_cuts, sep = ", ")
  dir_bin = head(factor(dir_intervals, labels = paste0("(", factor_labs, "]")),
                 -n_directions)
  
  
  ## Create a factor variable for wind speed intervals
  speed_cuts <- 3

      speed_cuts = c(0, speed_cuts)
    
    if (tail(speed_cuts, 1) < max(speed, na.rm = TRUE))
      speed_cuts = c(speed_cuts, max(speed, na.rm = TRUE))
    spd_bin = cut(speed, speed_cuts)
  }

  if (length(spd_cols) != length(levels(spd_bin)))
    spd_bin = cut_interval(speed, length(spd_cols))
  
  ## Create the dataframe suitable for plotting
    ggplot_df = as.data.frame(table(dir_bin, spd_bin, facet, facet_row))
    ggplot_df$proportion = unlist(by(ggplot_df$Freq, ggplot_df$facet,
                                     function(x) x / sum(x)),
                                  use.names = FALSE)
    ggplot_df$proportion <- ggplot_df$proportion * 3
=======
```{r wind-rose-5min, out.width = '100%'}
facet <- m$field
facet_row <- m$time_slice
speed <- m$wind_speed
direction <- m$wind_degrees
calm_wind <- 0
n_speeds <- 5
variable_wind <- 990
n_directions <- 12
>>>>>>> Stashed changes


  ggplot(data = ggplot_df,
                      aes_string(x = "dir_bin",
                                 fill = "spd_bin",
                                 y = "proportion")) +
    geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
    scale_x_discrete(
      breaks = levels(ggplot_df$dir_bin)[seq(1, n_directions,
                                             n_directions / 4)],
      labels = c("N", "E", "S", "W"),
      drop = FALSE
    ) +
    scale_fill_manual(name = "Wind speed (m/s)", values = spd_cols) +
    coord_polar(start = 2 * pi - pi / n_directions) +
    scale_y_continuous(labels = percent_format()) +
    theme(axis.title = element_blank()) +
    facet_grid(facet ~ facet_row)

wr
```

```{r save-wr, echo=FALSE, eval=TRUE}
ggsave(
  wr,
  file = "plots_manuscript/wind_rose_6_tr.png",
  width = 5,
  height = 8,
  units = "cm",
  dpi = 300,
  scale = 3
)
ggsave(
  wr,
  file = "plots_manuscript/wind_rose_6_tr.eps",
  device = cairo_ps,
  fallback_resolution = 600,
  width = 5,
  height = 8
)
```

## Wind Speed per Time Slice per Field

```{r wind_speed_per_slice, eval=TRUE, out.width = '100%'}
load(file = here("data/meteo_data.Rdata"))

ts_90 <- filter(m, time_slice == 90)
ts_180 <- filter(m, time_slice == 180)

ws <- ggplot(data = m, aes(x = minute_lap, y = wind_speed)) +
  geom_vline(xintercept = c(90, 180), lty = "dashed") +
  geom_text(aes(label = "90",
                x = 90,
                y = 13),
            angle = 90,
            vjust = 1) +
  geom_text(aes(label = "180",
                x = 180,
                y = 13),
            angle = 90,
            vjust = 1) +
  ylim(0, 15) +
  geom_point(size = 1.5, aes(colour = time_slice)) +
  geom_smooth(method = "gam") +
  scale_colour_viridis_c(breaks = c(90, 180, 270),
                         labels = c("90", "180", "270")) +
  guides(colour = guide_legend("Time Slice")) +
  ylab("Wind Speed (m/s)") +
  xlab("Time (5 min.)") +
  facet_wrap(. ~ field) +
  ggtitle(label = "Wind Speed")

ws
```

```{r save-ws, echo=FALSE, eval=TRUE}
ggsave(
  ws,
  file = "plots_manuscript/wind_speed_5min.png",
  width = 8,
  height = 5,
  units = "cm",
  dpi = 300,
  scale = 3
)
ggsave(
  ws,
  file = "plots_manuscript/wind_speed_5min.eps",
  device = cairo_ps,
  fallback_resolution = 600,
  width = 8,
  height = 5
)
```

## Plot smoothed spore dispersal

Plot spore dispersal as points with median values represented as red points with a line connecting the median values for each respective field.

```{r smoothed-dispersal}
field_names <- as_labeller(c(`1` = "Field 1",
                             `2` = "Field 2",
                             `3` = "Field 3",
                             `4` = "Field 4",
                             `5` = "Field 5",
                             `6` = "Field 6"))
smoothed_dispersal <- mod_dat %>%
  ggplot(aes(x = distance_m, y = n_spore)) +
  geom_point(size = 1.5) +
  stat_summary(fun = "median",
               geom = "line",
               na.rm = TRUE) +
  stat_summary(
    fun = "median",
    colour = "red",
    size = 2.5,
    geom = "point"
  ) +
  scale_x_continuous(guide = guide_axis(n.dodge = 2)) +
  facet_grid(. ~ field, labeller = field_names) +
  labs(col = "Prevailing wind",
       x = "Distance from harvesting field (m)",
       y = "Smut spores (n) at trap")
smoothed_dispersal
```

```{r save-spore-gradient, eval=TRUE, echo=FALSE}
ggsave(
  last_plot(),
  file = "plots_manuscript/spore_gradient.png",
  width = 8,
  height = 5,
  units = "cm",
  dpi = 300,
  scale = 3
)
ggsave(
  last_plot(),
  file = "plots_manuscript/spore_gradient.eps",
  device = cairo_ps,
  fallback_resolution = 600,
  width = 8,
  height = 5
)
```

## Spore type frequency

```{r load-spore-type}
dat_spore_type <- read_csv(here("data/spore_type.csv"))
```

```{r check-spore-type}
head(dat_spore_type)
```

```{r plot-spore-type}
dat_spore_type %>%
  group_by(meter, spore_nuclei, .drop = FALSE) %>%
  summarise(count = mean(count)) %>%
  mutate(percent = count / sum(count)
         ) -> dat_spore_type_sum

spore_type_plot <- dat_spore_type_sum %>%
  ggplot(aes(
    x = meter,
    y = percent,
    fill = factor(spore_nuclei)
  )) +
  geom_bar(position = position_fill(reverse = T),
           stat = "identity",
           alpha = .9) +
  labs(y = "Proportion",
       x = "Distance from harvesting field (m)",
       fill = "Spore type") +
  # scale_fill_viridis_d(direction = -1) +
  geom_text(
    data = dat_spore_type_sum %>%
      mutate(spore_nuclei = fct_rev(factor(spore_nuclei))),
    aes(label = round(percent, 2)),
    position = position_stack(vjust = 0.5),
    col = "white",
    fontface = "bold",
    size = 3
  ) +
  theme(#axis.title.y=element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) +
  guides(fill = "none")
```

## Spore size

```{r load-spore-size}
dat_spore_size <- read_csv(here("data/spore_size.csv"))
```

```{r inspect-spore-size}
head(dat_spore_size)
```

```{r spore-size-ridges}
med_spore_size <- dat_spore_size %>%
  filter(image_id != 34) %>%
  mutate(d_mean = (d1 + d2) / 2, 
         spore_type = as.factor(spore_type)) %>%
  group_by(spore_type) %>%
  summarize(mean=mean(d_mean), 
            sd=sd(d_mean, na.rm = T)) %>% 
  mutate(across(where(is.numeric), round, 1))

med_spore_size
```


```{r spore-size-ridges plot}
spore_size_plot <- dat_spore_size %>%
  filter(image_id != 34) %>%
  mutate(mean = (d1 + d2) / 2) %>%
  ggplot(aes(x = mean)) +
  geom_density(aes(fill = factor(spore_type)), size = .2, alpha = .6) +
  expand_limits(x = c(10, 50), y = c(0, .5)) +
  # scale_fill_viridis_d(direction = -1) +
  geom_vline(data = med_spore_size, aes(xintercept = mean), size = .2) +
  geom_label(
    data = med_spore_size,
    label.size = NA,
    aes(
      x = mean,
      label = paste0(mean, "\n(", sd, ")"),
      y = .48
    ),
    size = 2.54
  ) +
  labs(x = expression(paste("Mean diameter (", mu, m, ")")),
       y = "Density",
       fill = "Spore type")
```

## Create a multipanel figure of spore size details

```{r create-single-figure, fig.width = 7, fig.height = 5,   out.width = '100%'}
library(patchwork)
spore_size_plot + spore_type_plot +
  plot_layout(guides = "collect") &
  theme(legend.position = 'bottom') &
  scale_fill_viridis_d(direction = -1) &
  plot_annotation(tag_levels = 'A')
```

```{r save-spore-type-size, eval=TRUE, echo=FALSE}
ggsave(
  last_plot(),
  file = "plots_manuscript/spore_size_type.png",
  width = 8,
  height = 5,
  units = "cm",
  dpi = 300,
  scale = 3
)
ggsave(
  last_plot(),
  file = "plots_manuscript/spore_size_type.eps",
  device = cairo_ps,
  fallback_resolution = 600,
  width = 8,
  height = 5
)
```
