---
title: "Visualise data"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float:
      collapsed: no
      smooth_scroll: yes
    code_folding: hide
    toc_depth: 2
    css: style.css
---

```{r markdown-settings, echo=FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 7,
  message = FALSE,
  warning = FALSE
)
```

## Load libraries

```{r session settings}
library("tidyverse")
library("ggpubr")
library("here")
library("clifro")

theme_set(theme_pubclean())
```

## Load data

```{r load-data}
load(here("data/dat.Rdata"))
load(here("data/mod_dat.Rdata"))
```

## Plot spore dispersal density

### Plot spore densities

Create a density plot of the observed spore dispersal values.

```{r spore-kernel-density}
ggplot(mod_dat, aes(y = spore_cm2)) +
  geom_density()
```

Create a scatter plot and use `stat_smooth()` to fit a line.

```{r spore-dot-density-by-distance}
ggplot(mod_dat, aes(x = distance_m, y = spore_cm2)) +
  geom_point() +
  scale_fill_viridis_c() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 3)) +
  labs(y = expression(Spores/cm^{2}),
       x = "Distance (m)") +
  facet_grid(time_slice ~ .)
```

Create a violin plot with each time as a colour.

```{r spore-violin-density-by-distance}
ggplot(mod_dat, aes(x = as.factor(distance_m), y = spore_cm2)) +
  geom_violin(aes(fill = as.factor(time_slice),
                  colour = as.factor(time_slice))) +
  scale_fill_viridis_d("Time (minutes)") +
  scale_colour_viridis_d("Time (minutes)") +
  labs(y = expression(Spores/cm^{2}),
       x = "Distance (m)")
```

#### Check windspeed effect

Create a hexbin plot to display the number of spores by wind speed and fit a smoothed line to these variables.

```{r spore-windspeed}
ggplot(data = mod_dat, aes(x = wind_speed, y = spore_cm2)) +
  scale_fill_viridis_c() +
  geom_hex() +
  geom_smooth() +
  labs(y = expression(Spores / cm ^ {
    2
  }),
  x = "Wind Speed (m/s)") +
  theme(
    legend.position = "top",
    legend.background = element_blank(),
    legend.key = element_blank()
  ) +
  guides(fill = guide_legend(title.position = "top",
                             title.hjust = 0.5))
```

#### Polar coordinates

Create polar coordinate plots that display spore density in each cardinal direction.

```{r plot-spores}
ggplot(data = mod_dat) +
  aes(
    x = trap_degrees,
    y = distance_m,
    colour = spore_cm2,
    size = spore_cm2
  ) +
  facet_grid(time_slice ~ field) +
  coord_polar(theta = "x",
              start = 0,
              direction = 1) +
  geom_count(alpha = 0.55) +
  scale_colour_viridis_c(
    direction = -1,
    name = "Smut spores/m²",
    guide = "legend",
    breaks = c(10, 100, 200, 300),
    begin = 0,
    end = 0.8
  ) +
  scale_size(
    range = c(1, 8),
    name = "Smut spores/m²",
    breaks = c(10, 100, 200, 300)
  ) +
  scale_x_continuous(
    breaks = c(0, 90, 180, 270),
    expand = c(0, 0),
    limits = c(0, 360),
    labels = c("N", "E", "S", "W"),
    sec.axis = sec_axis(
      ~ . ,
      name = "Field",
      breaks = NULL,
      labels = NULL
    )
  ) +
  scale_y_continuous(
    breaks = c(0, 100, 200, 300, 400),
    limits = c(0, 400),
    sec.axis = sec_axis(
      ~ . ,
      name = "Minutes since harvest beginning",
      breaks = NULL,
      labels = NULL
    )
  ) +
  ylab("Distance (m)") +
  xlab("") +
  theme(
    legend.position = "top",
    legend.background = element_blank(),
    legend.key = element_blank()
  ) +
  guides(color = guide_legend(title.position = "top",
                              title.hjust = 0.5))
```

```{r eval=FALSE}
ggsave(
  last_plot(),
  file = "plots/spores_cardinal.png",
  width = 20,
  height = 10,
  units = "cm",
  dpi = 600,
  scale = 1.5
)
```

## Plot wind data

### Wind roses

Using the `wind_rose()` from _clifro_, create windrose plots to visualise windspeeds and directions during the harvest and spread events.

```{r wind-rose}
wind_rose <-
  with(
    mod_dat,
    windrose(
      speed = wind_speed,
      direction = wind_degrees,
      facet = field,
      n_speeds = 5,
      n_col = 3
    )
  )

wind_rose +
  scale_fill_viridis_d(
    name = "Wind Speed (m/s)",
    direction = -1,
    begin = 0,
    end = .8
  ) +
  labs(x = "", y = "") +
  theme(legend.position = 'top') +
  guides(fill = guide_legend(
    title.position = "top",
    title.hjust = 0.5,
    keywidth = unit(1, "lines"),
    keyheight = unit(1, "lines")
  )) +
  theme_pubclean()
```

```{r}
ggsave(
  last_plot(),
  file = "plots/wind_rose_6_tr.png",
  width = 20,
  height = 10,
  units = "cm",
  dpi = 600,
  scale = 1.5
)
```

## Plot smoothed spore dispersal

Plot spore dispersal with smoothed lines for spore traps that are classified as "upwind", "downwind" or "lateral".

```{r smoothed-dispersal}
mod_dat %>%
  ggplot(aes(distance_m, spore_cm2, col = downwind)) +
  geom_point() +
  geom_smooth(se = F) +
  facet_grid(field ~ time_slice) +
  scale_x_continuous(
    sec.axis = sec_axis(
      ~ . ,
      name = "Minutes since harvest beginning",
      breaks = NULL,
      labels = NULL
    )
  ) +
  scale_y_continuous(
    sec.axis = sec_axis(
      ~ . ,
      name = "Field",
      breaks = NULL,
      labels = NULL
    )
  ) +
  scale_colour_viridis_d() +
  labs(col = "Prevailing wind",
       x = "Distance from harvesting field (m)",
       y = "Smut spores / m² at trap") +
  theme(
    legend.position = "top",
    legend.justification = "center",
    legend.direction = "horizontal",
    legend.background = element_blank(),
    legend.key = element_blank()
  ) +
  guides(colour = guide_legend(title.position = "top",
                               title.hjust = 0.5))
```

```{r, eval=FALSE}
ggsave(
  last_plot(),
  file = "plots/spore_gradient.png",
  w = 6,
  h = 4,
  scale = 1.2
)
```
