---
title: "Import data"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float:
      collapsed: no
      smooth_scroll: yes
    code_folding: hide
    toc_depth: 2
    css: style.css
editor_options: 
  chunk_output_type: console
---

```{r markdown-settings, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 7,
  message = FALSE,
  warning = FALSE
)
```

## Load libraries and function

A specific function for circular averaging the wind direction is necessary to summarise these data.
This function will be loaded from `./R/circular_averaging.R` after loading the necessary libraries.

```{r libraries}
library("tidyverse")
library("here")
library("geometry")

# load circular.averaging()
source(here("./R/circular_averaging.R"))
```

## Import compiled dataset from local files.

### Spore count data

Import spore trap data and inspect.

```{r spore-data}
spore_data <- read_csv(
  here("data/spore_data.csv"),
  col_types = list(field = col_factor(),
                   trap_coord = col_factor()
  ))
```

Summarise the spore data into the mean number of spores per centimeter squared combining the traps' samples and sub-subsamples.
We'll keep the `field` col to join the weather data and the `trap_coord` col here for joining with another data frame that has the cardinal directions in it for further analysis.

```{r summarise-spore-data}
spore_data <- 
  spore_data %>%
  group_by(field, trap_coord, distance_m, time_min) %>%
  summarise(spore_cm2 = mean(n_spore) / 1.2)
spore_data
```

### Weather data

Import weather data, filter unused data, convert to SI units and inspect the resulting data.

```{r meteo-data}
meteo_data <- read_csv(here("data/wind.csv"),
                       col_types = list(field = col_factor()))

# some sites go past 270 minutes, remove those data.
meteo_data <- 
  meteo_data %>% 
  filter(minute_lap <= 270) %>% 
  mutate(wind_speed = wind_speed * 0.2777778)

meteo_data
```

### Data prep for analysis

#### Add cardinal directions

We need to convert cardinal directions to degrees for plotting and modeling purposes (adapted from Adam's code for chickpea ascochyta blight spore dispersal).

```{r cardinal-directions}
coord_deg <- tibble(
  dir = c(
    "N",
    "NNE",
    "NE",
    "ENE",
    "E",
    "ESE",
    "SE",
    "SSE",
    "S",
    "SSW",
    "SW",
    "WSW",
    "W",
    "WNW",
    "NW",
    "NNW"
  ),
  degrees = seq(0, 337.5 , by = 22.5),
)
```

Create a `spore_data` object that has spore and trap degrees (location as a cardinal direction).

```{r add-spore-trap-location}
spore_data <- 
  spore_data %>%
  left_join(coord_deg, by = c("trap_coord" = "dir")) %>%
  rename(trap_degrees = degrees)
```

Create a `meteo_data` object that has the wind direction in degrees.

```{r add-wind-direction-degrees}
meteo_data <- 
  meteo_data %>%
  left_join(coord_deg, by = c("wind_dir" = "dir")) %>%
  rename(wind_degrees = degrees)
```

#### Summarise the weather data into 90, 180 and 270 minute interval bins

The weather data are summarised from 0 to 90, 0 to 180 and 0 to 270 minutes to align with the spore trapping data.

This is where the `circular.averaging()` is needed since we cannot simply take the `mean` value of wind direction in degrees here.

```{r bin-time-slices}
m <-
  meteo_data %>%
  mutate(time_slice =
             case_when(
               minute_lap <= 90 ~ 90,
               minute_lap > 90 & minute_lap <= 180 ~ 180,
               TRUE ~ 270
             )
           )
m
```

<<<<<<< HEAD
```{r meteo_data}
=======

```{r bin-time-slices-rh, eval=TRUE}
rh_90 <-
  m %>%
  filter(time_slice == 90) %>%
  group_by(field) %>%
  summarise(rh = mean(rel_hum))

rh_180 <-
  m %>%
  filter(time_slice == 90 | time_slice == 180) %>%
  group_by(field) %>%
  summarise(rh = mean(rel_hum))

rh_270 <-
  m %>%
  group_by(field) %>%
  summarise(rh = mean(rel_hum))

rh <-
  bind_rows(list(
    "90" = rh_90,
    "180" = rh_180,
    "270" = rh_270
  ), .id = "time_slice")
```


```{r bin-time-slices-temp, eval=TRUE}
temp_90 <-
  m %>%
  filter(time_slice == 90) %>%
  group_by(field) %>%
  summarise(temp = mean(temp))

temp_180 <-
  m %>%
  filter(time_slice == 90 | time_slice == 180) %>%
  group_by(field) %>%
  summarise(temp = mean(temp))

temp_270 <-
  m %>%
  group_by(field) %>%
  summarise(temp = mean(temp))

temp <-
  bind_rows(list(
    "90" = temp_90,
    "180" = temp_180,
    "270" = temp_270
  ), .id = "time_slice")
```


```{r bin-time-slices-wind, eval=TRUE}
wind_speed_90 <-
  m %>%
  filter(time_slice == 90) %>%
  group_by(field) %>%
  summarise(wind_speed = mean(wind_speed))

wind_speed_180 <-
  m %>%
  filter(time_slice == 90 | time_slice == 180) %>%
  group_by(field) %>%
  summarise(wind_speed = mean(wind_speed))

wind_speed_270 <-
  m %>%
  group_by(field) %>%
  summarise(wind_speed = mean(wind_speed))

wind_speed <-
  bind_rows(list(
    "90" = wind_speed_90,
    "180" = wind_speed_180,
    "270" = wind_speed_270
  ), .id = "time_slice")
```


```{r bin-time-slices-wind_degrees, eval=TRUE}
wind_degrees_90 <-
  m %>%
  filter(time_slice == 90) %>%
  group_by(field) %>%
  summarise(wind_degrees = circular.averaging(wind_degrees))

wind_degrees_180 <-
  m %>%
  filter(time_slice == 90 | time_slice == 180) %>%
  group_by(field) %>%
  summarise(wind_degrees = circular.averaging(wind_degrees))

wind_degrees_270 <-
  m %>%
  group_by(field) %>%
  summarise(wind_degrees = circular.averaging(wind_degrees))

wind_degrees <-
  bind_rows(list(
    "90" = wind_degrees_90,
    "180" = wind_degrees_180,
    "270" = wind_degrees_270
  ), .id = "time_slice")

meteo_data <-
  rh %>% 
  left_join(temp, by = c("field", "time_slice")) %>% 
  left_join(wind_speed, by = c("field", "time_slice")) %>% 
  left_join(wind_degrees, by = c("field", "time_slice")) %>% 
  mutate(time_slice = as.numeric(time_slice)) %>% 
  arrange(field, time_slice)
```

```{r meteo_data_JUAN}
>>>>>>> ce84285317605d7534788d7b41df2fe9ccc6b76c
meteo_data <- m %>%
  mutate(time_slice = 270) %>%  
  rbind(mutate(filter(., minute_lap <= 180), time_slice = 180)) %>% 
  rbind(mutate(filter(., minute_lap <= 90), time_slice = 90)) %>% 
  group_by(field, time_slice) %>%
  summarize(rh = mean(rel_hum), 
            temp = mean(temp),
            wind_speed = mean(wind_speed), 
            wind_degrees = circular.averaging(wind_degrees))
```

## Merge the spore data with the weather data

Use `left_join` to merge the raw spore data, `spore_raw` and weather data, `meteo_data` on `field` and `time_slice`/`time_min`.
Then add columns for the x and y locations on a Cartesian grid of the traps.

```{r merge-spore-weather}
mod_dat <- left_join(meteo_data,
                     spore_data,
                     by = c("field" = "field",
                            "time_slice" = "time_min")) %>%
  mutate(degree_dif = abs(trap_degrees - wind_degrees)) %>%
  mutate(time_slice = as.integer(time_slice))
```

## Create Cartesian coords

```{r Cartesian-coords}
# degrees need to be converted to radians for theta
xy <- data.frame(pol2cart(theta = mod_dat$trap_degrees * 0.01745329,
                          r = mod_dat$distance_m))

mod_dat <- bind_cols(mod_dat, xy)
```


```{r Cartesian-coords-check, eval=FALSE}
# mod_dat %>% view

coords_df <- mod_dat %>% 
  ungroup() %>% 
  distinct(trap_degrees, distance_m, x, y) %>% 
  mutate_at(vars(x,y), list(~ round(., 1))) %>% 
  arrange(trap_degrees) # %>% print(n = Inf)

coords_df %>%   
  ggplot()+
  aes(x=y, y=x)+
  geom_point() +
  geom_text(aes(label=paste0(trap_degrees,",", distance_m), vjust=1))+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0) +
  coord_equal()
```

Save data object for fitting models.

```{r save-mod-dat, eval=TRUE}
save(mod_dat, file = here("data/mod_dat.Rdata"))
```
