---
title: "Fit GAM model of spore counts"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float:
      collapsed: no
      smooth_scroll: yes
    code_folding: hide
    toc_depth: 2
    css: style.css
editor_options: 
  chunk_output_type: inline
---

```{r markdown-settings, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 7,
  message = FALSE,
  warning = FALSE
)
```

## Load libraries

```{r session settings}
library("tidyverse")
library("broom")
library("ggplot2")
library("ggpubr")
library("gratia")
library("here")
library("mgcv")

theme_set(theme_pubclean())
```

## Load data

```{r load-data}
load(here("data/mod_dat.Rdata"))
```

## Inspect the mod_dat object

```{r inspect-mod-dat}
mod_dat %>% # str
  distinct(field, trap_coord, .keep_all = TRUE) %>%
  select(trap_coord,
         trap_degrees,
         degree_dif,
         wind_degrees,
         wind_speed,
         distance_m) %>%
  print(n = Inf)
```

Based on the use of AIC to evaluate model fit to the data, the following model best fit the data with smoothed terms for x and y, the Cartesian coordinates of the spore traps placed around the harvest fields, `x` and `y`; the smoothed time slice variables, `time_slice`; the wind direction in relation to the traps' placement, `degree_dif`; the traps' distance from the harvest field, `distance_m`; and the field and traps themselves as a random effects, `field` and `xy`.

```{r m1}
mod_dat$xy <- as.factor(paste(mod_dat$x, mod_dat$y))

m1 <- gam(
  spore_cm2 ~ s(y, x, k = 20) +
    s(time_slice, k = 3) +
    s(degree_dif, k = 72) +
    s(distance_m, k = 4) +
    s(field, xy, bs = "re"),
  data = mod_dat,
  select = TRUE,
  method = "REML",
  family = "tw"
)

m1
```

## Inspect Model m1

### Check model

```{r gam-check-m1}
gam.check(m1)
summary(m1)
```

The model fully converged after 10 iterations.
All of the predictive factors included were significant at _P_ > 0.05 and the smoothed predictors indicate a good fit with residuals being randomly distributed and enough basis functions for them to be properly fit.
The Q-Q plot generally falls along the line indicating a sufficient fit.
There is a bit of a pattern in the residual vs linear predictors but the histogram of residuals is a nice bell-shape.
The response versus fitted values clusters nicely around the 1:1 line.
In all, this model looks to explain the spore dispersal data very well.

#### Check for autocorrelation

Because the data represent spatially and temporally correlated data, it's best to check for any autocorrelation in the residuals.

```{r gam-autocorrelation}
acf(residuals(m1))
pacf(residuals(m1))
```

This looks OK.
There is no major pattern discernible in the ACF or PACF plots that show cause for concern.

### Summarise and visualise model m1

```{r appraise-m1}
draw(m1)
```

```{r, eval=FALSE}
ggsave(
  last_plot(),
  file = "plots_manuscript/model_summary.jpg",
  width = 8,
  height = 5,
  units = "cm",
  dpi = 300,
  scale = 3
)
```

## Predict spore dispersal

```{r predict}
dist <- seq(0, 400, by = 100)

newd <- expand_grid(
  time_slice = c(90, 180, 270),
  degree_dif = 1:360,
  distance_m = dist,
  wind_speed = 0:30,
  x = dist,
  y = dist,
  xy = paste(dist, -dist),
  field = 1)
  
newd$pred_spore_cm2 <- predict(m1, newd)
```

## Visualise spore dispersal predictions

Visualise the spore dispersal predictions for traps at 0m, 100m, 200m, 300m, 400m around the harvest field for sampling times at 90, 180 and 270 minutes after harvest started.

```{r graph-predictions}
newd %>%
  mutate(pred_spore_cm2 = replace(pred_spore_cm2, pred_spore_cm2 < 0, 0)) %>%
  ggplot(aes(x = as.factor(distance_m), y = pred_spore_cm2)) +
  geom_violin() +
  labs(x = "Distance from harvesting field (m)",
       y = "Predicted Spores / cm²"
  )
```

```{r save-prediction-graph, eval=TRUE}
ggsave(
  last_plot(),
  file = "plots_manuscript/spore_prediction.png",
  w = 5,
  h = 2,
  scale = 1.2
)
```
