---
title: "Fit GAM model of spore counts"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float:
      collapsed: no
      smooth_scroll: yes
    code_folding: hide
    toc_depth: 2
    css: style.css
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

```{r markdown-settings, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(
  out.width = '60%',
  fig.width = 7,
  fig.height = 7,
  message = FALSE,
  warning = FALSE
)
```

## Load libraries

```{r session settings}
library("tidyverse")
library("ggpubr")
library("gratia")
library("here")
library("mgcv")
library("DHARMa")
theme_set(theme_pubclean())
```

## Spore proportion chi-squared analysis

Check if the spore type, number of nuclei, affects the dispersal.

Using the default `chisq.test()` will result in a warning due to small expected values meaning that approximations of p may not be correct.
`
Warning message:
In chisq.test(.) : Chi-squared approximation may be incorrect
`
Therefore, we will use Monte Carlo simulation to simulate p values.

(Manuscript's section `Results/3.2`)

```{r x-sq}
set.seed(3)
x <- read_csv(here("data/spore_type.csv"),
              show_col_types = FALSE) %>%
  group_by(meter, spore_nuclei) %>%
  summarise(n = sum(count), .groups = 'drop') %>%
  spread(spore_nuclei, n) %>%
  column_to_rownames('meter') %>%
  chisq.test(simulate.p.value = TRUE)
```

The major spore type spread expected was the individual spore having a smaller spore size (18.8 µm); however, no statistical differences were recorded in spore type proportion at 100 m to 400 m downwind (P = `r x$p.value`, X-squared = `r x$statistic`).

## Spore dispersal GAM Model

(Manuscript's section `Results/3.3`)

### Load data

```{r load-data}
load(here("data/mod_dat.Rdata"))

sdp_dat <- read_csv(
  here("data/sdp_data.csv"),
  show_col_types = FALSE,
  col_types = list(field = col_factor())
) %>%
  arrange(field) %>%
  select(field, inc, SDP, DSI) %>%
  mutate(field = as.factor(field))

# summarise SDP index by field and create a binomial variable w/ cutoff @ 5 ----

ps_inc <-
  sdp_dat %>%
  group_by(field) %>%
  summarise(SDP = mean(SDP))

mod_dat <- left_join(mod_dat, ps_inc, by = c("field"))
```

### Inspect the mod_dat object

```{r inspect-mod-dat}
mod_dat %>% # str
  distinct(field, trap_coord, .keep_all = TRUE) %>%
  select(
    trap_coord,
    trap_degrees,
    degree_dif,
    wind_degrees,
    wind_speed,
    distance_m,
    SDP,
    n_spore
  ) %>%
  print(n = Inf)
```

Based on the use of AIC to evaluate model fit to the data, the following GAM best fit the data with the smoothed time slice variables, s(`time_slice`); a smoothed term for the the distance of the traps from the field and wind speed, `s(distance_m, wind_speed)`; a smoothed term for the severely damaged pods, `s(SDP)` and the field and traps themselves as a smoothed term random effect, `s(field`, `xy`) using a negative binomial family, `nb()`.

```{r fit-m1, eval=TRUE}
# create a factor for the trap locations in Cartesian coords to be used as a random effect
mod_dat$xy <- as.factor(paste(mod_dat$x, mod_dat$y))

m1 <- gam(
  n_spore ~ s(time_slice, k = 3) +
    s(distance_m, wind_speed, k = 50) +
    s(SDP, k = 6) +
    s(field, xy, bs = "re"),
  data = mod_dat,
  select = TRUE,
  method = "REML",
  family = nb()
)
```

### Inspect Model m1

```{r inspect-m1}
summary(m1)
```

#### Check model m1

```{r gam-check-m1}
gam.check(m1)

summary(m1)

vis.gam(
  m1,
  theta = 50,
  view = c("time_slice", "SDP"),
  ticktype = "detailed",
  main = "Time slice (minutes) and severely damaged pods (SDP)"
)

vis.gam(
  m1,
  view = c("distance_m", "wind_speed"),
  theta = 50,
  ticktype = "detailed",
  main = "Trap distance from field (m) and wind speed (m/s)"
)

plot(m1)

# check residuals using DHARMa
simulateResiduals(m1, plot = TRUE)
```

The model fully converged after 11 iterations.
All of the predictive factors included were significant at *P* \> 0.1 and the model had an adjusted R2 value of 0.726 and explains 89.2% of the null deviance.

The `gam.check()` residuals show a slight pattern in the Residuals vs. linear predicted values, but this is to be expected with the zero values and negative binomial family, `nb()`, that was used to fit the model. The QQ-plot appears to be reasonable as does the histogram of residuals, and Response vs Fitted Values.
To further check the model's fitness, there is a slight pattern in the DHARMa residuals, but nothing to worry about too much in a small data set like this while this QQ-plot and the associated tests all pass.

In all, this model looks to explain the spore dispersal data well enough for this data set.

#### Check for autocorrelation

Because the data represent spatially and temporally correlated data, it's best to check for any autocorrelation in the residuals.

```{r gam-autocorrelation}
acf(residuals(m1))
pacf(residuals(m1))
```

This looks OK.
There is no major pattern discernible in the ACF or PACF plots that show cause for concern.

```{r save-m1-summary, eval=FALSE, echo=FALSE}
appraise_m1 <- appraise(m1)
ggsave(
  apraise_m1,
  file = here("plots_manuscript/model_summary.png"),
  width = 8,
  height = 5,
  units = "cm",
  dpi = 300,
  scale = 3
)
ggsave(
  appraise_m1,
  file = here("plots_manuscript/model_summary.eps"),
  device = cairo_ps,
  fallback_resolution = 600,
  width = 8,
  height = 5
)
```

### Predict spore dispersal and visualise spore dispersal predictions

```{r predict, eval=TRUE}
dist <- seq(100, 400, by = 100)

newd <- expand_grid(
  wind_speed = 1:10,
  SDP = seq(from = 0.5, to = 10, by = 0.5),
  time_slice = unique(mod_dat$time_slice),
  distance_m = unique(mod_dat$distance_m),
  xy = unique(mod_dat$xy),
  field = 1
)

newd$pred_n_spore <-
  predict(
    object = m1,
    newd,
    type = "response",
    exclude = c("field", "xy")
  )
```

Visualize the spore dispersal predictions for traps at 100 m, 200 m, 300 m, 400 m around the harvest field for sampling times at 90, 180 and 270 minutes after harvest started.

```{r graph-predictions}
predictions <-
  newd %>%
  ggplot(aes(x = as.factor(distance_m), y = pred_n_spore)) +
  geom_boxplot() +
  labs(x = "Distance from harvesting field (m)",
       y = "Predicted Spores (n)")

predictions
```

```{r save-prediction-graph, eval=FALSE}
ggsave(
  predictions,
  file = "plots_manuscript/spore_prediction.png",
  w = 5,
  h = 2,
  scale = 1.2
)
```
