---
title: "Fit models"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float:
      collapsed: no
      smooth_scroll: yes
    code_folding: hide
    toc_depth: 2
    css: style.css
---

```{r markdown-settings, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 7,
  message = FALSE,
  warning = FALSE
)
```

## Load libraries

```{r session settings}
library("tidyverse")
library("broom")
library("ggplot2")
library("ggpubr")
library("gratia")
library("here")
library("mgcv")

theme_set(theme_pubclean())
```

## Load data

```{r load-data}
load(here("data/dat.Rdata"))
load(here("data/mod_dat.Rdata"))

# add a factor col for the spore traps in each field respectively
mod_dat$trap_id <-
  as.factor(paste0(mod_dat$field, mod_dat$trap_coord, mod_dat$distance_m))
mod_dat$sub_sample <- as.factor(mod_dat$sub_sample)
```

## Inspect the mod_dat object

```{r inspect-mod-dat}
mod_dat %>%
  distinct(field, trap_coord, .keep_all = TRUE) %>%
  select(trap_coord,
         trap_degrees,
         wind_dir,
         wind_degrees,
         degree_dif,
         downwind,
         trap_id) %>%
  print(n = Inf)
```

Because of the number of zeros in the data, we will use a [Tweedie family](https://en.wikipedia.org/wiki/Tweedie_distribution) for the models to help account for this, _i.e._ in the model functions you will see `family =  "tw"` for this.

### Model m1

* Response: `spore_cm2`
* Linear predictors: `distance_m:time_min`
* Random effects: `field`

```{r m1}
m1 <- gam(
  spore_cm2 ~ distance_m:time_min +
    s(field, bs = "re"),
  data = mod_dat,
  method = "REML",
  family = "tw"
)
m1
```

### Model m2

* Response: `spore_cm2`
* Linear predictors: `distance_m:time_min`
* Smoothed predictors: `downwind`
* Random effects: `field`

```{r m2}
m2 <- gam(
  spore_cm2 ~ distance_m:time_min +
    s(field, bs = "re") +
    s(downwind, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m2
```

### Model m3

* Response: `spore_cm2`
* Linear predictors: `distance_m:time_min`
* Random effects: `field`, `downwind`, `wind_speed`

```{r m3}
m3 <- gam(
  spore_cm2 ~ distance_m:time_min +
    s(wind_speed, bs = "re") +
    s(field, bs = "re") +
    s(downwind, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m3
```

### Model m4

* Response: `spore_cm2`
* Linear predictors: `time_min`, `downwind`, `wind_speed`
* Smoothed predictors: `distance_m`
* Random effects: `field`

```{r m4}
m4 <- gam(
  spore_cm2 ~ time_min + downwind + wind_speed +
    s(distance_m, k = 3) +
    s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m4
```

### Model m5

* Response: `spore_cm2`
* Linear predictors: `time_min`, `downwind`, `wind_speed`
* Smoothed predictors: `distance_m`
* Random effects: `field`, `time_min`

```{r m5}
m5 <- gam(
  spore_cm2 ~ time_min + downwind + wind_speed +
    s(distance_m, k = 3) +
    s(time_min, bs = "re") +
    s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m5
```

### Model m6

* Response: `spore_cm2`
* Linear predictors: `time_min`, `downwind`
* Smoothed predictors: `distance_m`
* Random effects: `field`

```{r m6}
m6 <- gam(
  spore_cm2 ~ time_min + downwind + s(distance_m, k = 3) +
    s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m6
```

### Model m7

* Response: `spore_cm2`
* Linear predictors: `time_min`, `distance_m`, `downwind`
* Random effects: `field`

```{r m7}
m7 <- gam(
  spore_cm2 ~ time_min + distance_m + downwind + s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m7
```

### Model m8

* Response: `spore_cm2`
* Linear predictors: `time_min`, `distance_m`, `downwind`
* Smoothed predictors: `wind_speed`
* Random effects: `field`

```{r m8}
m8 <- gam(
  spore_cm2 ~ distance_m + downwind + time_min +
    s(wind_speed, k = 6) +
    s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m8
```

### Model m9

* Response: `spore_cm2`
* Linear predictors: `time_min`, `distance_m`, `downwind`, `smut_incidence`
* Smoothed predictors: `wind_speed`
* Random effects: `field`

```{r m9}
m9 <- gam(
  spore_cm2 ~ distance_m + downwind + time_min + `smut_incidence` +
    s(wind_speed, k = 6) +
    s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m9
```

### Model m10
* Response: `spore_cm2`
* Linear predictors: `time_min`, `distance_m`, `downwind`, `smut_incidence`
* Smoothed predictors: `wind_speed`
* Random effects: `trap_id`, `sub_sample`

```{r m10}
m10 <- gam(
  spore_cm2 ~ time_min + distance_m + downwind +
    s(wind_speed, k = 6, by = downwind) +
    s(trap_id, sub_sample, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m10
```

### Model m11
* Response: `spore_cm2`
* Linear predictors: `time_min`, `distance_m`, `downwind`, `smut_incidence`
* Smoothed predictors: `wind_speed`
* Random effects: `trap_id` and `sub_sample` grouped by `trap_id`

```{r m11}
m11 <- gam(
  spore_cm2 ~ time_min + distance_m + downwind +
    s(wind_speed, k = 6) +
    s(trap_id, bs = "re") +
    s(sub_sample, by = trap_id, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m11
```

### Model m12
* Response: `spore_cm2`
* Linear predictors: `time_min`, `distance_m`, `downwind`, `smut_incidence`
* Smoothed predictors: `wind_speed`
* Random effects: `trap_id`, `sub_sample`

```{r m12}
m12 <- gam(
  spore_cm2 ~ time_min + distance_m + downwind +
    s(wind_speed, k = 6) +
    s(trap_id, bs = "re") +
    s(sub_sample, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m12
```

## Model AIC

Check models for best fit using AIC.

```{r AIC}
models <- list(
  m1 = m1,
  m2 = m2,
  m3 = m3,
  m4 = m4,
  m5 = m5,
  m6 = m6,
  m7 = m7,
  m8 = m8,
  m9 = m9,
  m10 = m10,
  m11 = m11,
  m12 = m12
)
map_df(models, glance, .id = "model") %>%
  arrange(AIC, BIC)
```

Based on the AIC and BIC values, model m12 looks to be the best model for our data.

## Inspect Model m12

### Check model

```{r gam-check-m12}
gam.check(m12)
```

The model did fully converge after 5 iterations but wind speed is residuals are not randomly distributed.
The diagnostic plots show some issues with the fit, indicating that we really need more data to fit a proper model if we were interested in fully predicting spore dispersal, but for our purposes, explaining the dispersal, this model is good enough.
However, in the future, collecting more data so that `k` could be increased would be useful, but since this is the data that we have, this appears to be the best model that describes it.

### Summarise model

```{r appraise-m12}
summary(m12)
```

### Visualise the model

```{r visualise-m12}
draw(m12)
```

## Predict spore dispersal

```{r predict}
mod_dat %>% distinct(downwind) %>% pull()
newd <- expand_grid(
  time_min = c(90, 180, 270),
  downwind = c("downwind", "lateral", "upwind"),
  distance_m = seq(0, 400, 100),
  wind_speed = 0:30,
  trap_id = as.factor(apply(expand.grid(
    1, c("N", "S", "E", "W"), c(100, 200, 300, 400)
  ),
  1, paste, collapse = "")),
  sub_sample = as.factor(c(1, 2))
)
  
newd$pred_spore_cm2 <- predict(m12, newd)
```

## Visualise spore dispersal predictions

The GAM will predict spores to disperse at negative values, since this is not possible, we will set any values that are less than zero to zero in the model's output before plotting.

```{r graph-predictions, fig.cap="Predictions of spore dispersal (spores per meter squared) at 0m to 400m for three time points, 90 min, 180 min, 270 min after peanut harvest commenced."}
newd %>%
  mutate(pred_spore_cm2 = replace(pred_spore_cm2, pred_spore_cm2 < 0, 0)) %>%
  ggplot(aes(x = distance_m, y = pred_spore_cm2, col = downwind)) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), size = 1) +
  scale_fill_viridis_d("Prevailing wind") +
  scale_colour_viridis_d("Prevailing Wind") +
  facet_grid(. ~ time_min) +
  scale_x_continuous(sec.axis = sec_axis(
    ~ . ,
    name = "Minutes since harvest beginning",
    breaks = NULL,
    labels = NULL
  )) +
  labs(x = "Distance from harvesting field (m)",
       y = "Spores / m²") +
  theme(
    legend.position = "top",
    legend.justification = "center",
    legend.direction = "horizontal",
    legend.background = element_blank(),
    legend.key = element_blank()
  ) +
  guides(fill = guide_legend(title.position = "top",
                             title.hjust = 0.5))
```

```{r save-prediction-graph, eval=FALSE}
ggsave(
  last_plot(),
  file = "plots/spore_prediction.png",
  w = 5,
  h = 2,
  scale = 1.2
)
```



