---
title: "Fit models"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float:
      collapsed: no
      smooth_scroll: yes
    code_folding: hide
    toc_depth: 2
    css: style.css
---

```{r markdown-settings, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 7,
  message = FALSE,
  warning = FALSE
)
```

## Load libraries

```{r session settings}
library("tidyverse")
library("broom")
library("ggplot2")
library("ggpubr")
library("gratia")
library("here")
library("mgcv")

theme_set(theme_pubclean())
```

## Load data

```{r load-data}
load(here("data/dat.Rdata"))
load(here("data/mod_dat.Rdata"))
```

## Inspect the mod_dat data

```{r inspect-mod-dat}
mod_dat %>%
  distinct(field, trap_coord, .keep_all = TRUE) %>%
  select(trap_coord,
         trap_degrees,
         wind_dir,
         wind_degrees,
         degree_dif,
         downwind) %>%
  print(n = Inf)
```

Because of the number of zeros in the data, we will use a [Tweedie family](https://en.wikipedia.org/wiki/Tweedie_distribution) for the models to help account for this, _i.e._ in the model functions you will see `family =  "tw"` for this.

### Model 1

* Response: `spore_cm2`
* Linear predictors: `distance_m:time_min`
* Random effects: `field`

```{r m1}
m1 <- gam(
  spore_cm2 ~ distance_m:time_min +
    s(field, bs = "re"),
  data = mod_dat,
  method = "REML",
  family = "tw"
)
m1
```

### Model 2

* Response: `spore_cm2`
* Linear predictors: `distance_m:time_min`
* Smoothed predictors: `downwind`
* Random effects: `field`

```{r m2}
m2 <- gam(
  spore_cm2 ~ distance_m:time_min +
    s(field, bs = "re") +
    s(downwind, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m2
```

### Model 3

* Response: `spore_cm2`
* Linear predictors: `distance_m:time_min`
* Random effects: `field`, `downwind`, `wind_speed`

```{r m3}
m3 <- gam(
  spore_cm2 ~ distance_m:time_min +
    s(wind_speed, bs = "re") +
    s(field, bs = "re") +
    s(downwind, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m3
```

### Model 4

* Response: `spore_cm2`
* Linear predictors: `time_min`, `downwind`, `wind_speed`
* Smoothed predictors: `distance_m`
* Random effects: `field`

```{r m4}
m4 <- gam(
  spore_cm2 ~ time_min + downwind + wind_speed +
    s(distance_m, k = 3) +
    s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m4
```

### Model 4.1

* Response: `spore_cm2`
* Linear predictors: `time_min`, `downwind`, `wind_speed`
* Smoothed predictors: `distance_m`
* Random effects: `field`, `time_min`

```{r m4.1}
m4.1 <- gam(
  spore_cm2 ~ time_min + downwind + wind_speed +
    s(distance_m, k = 3) +
    s(time_min, bs = "re") +
    s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m4.1
```

### Model 5

* Response: `spore_cm2`
* Linear predictors: `time_min`, `downwind`
* Smoothed predictors: `distance_m`
* Random effects: `field`

```{r m5}
m5 <- gam(
  spore_cm2 ~ time_min + downwind +
    s(distance_m, k = 3) +
    s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m5
```

### Model 5.1



```{r m5.1}
m5.1 <- gam(
  spore_cm2 ~ time_min + downwind + distance_m +
    s(wind_speed, k = 6) +
    s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m5.1
```

### Model 6

* Response: `spore_cm2`
* Linear predictors: `time_min`, `distance_m`, `downwind`, `wind_speed`, `rel_hum`, `temp`, `wind_speed`, `smut_incidence`
* Smoothed predictors: `distance_m`
* Random effects: `field`, `time_min`

```{r m6}
m6 <- gam(
  spore_cm2 ~ distance_m + downwind + time_min +
    s(wind_speed, k = 6) +
    s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m6
```

## Model AIC

Check models for best fit using AIC.

```{r AIC}
models <- list(
  m1 = m1,
  m2 = m2,
  m3 = m3,
  m4 = m4,
  m4.1 = m4.1,
  m5 = m5,
  m5.1 = m5.1,
  m6 = m6
)
map_df(models, glance, .id = "model") %>%
  arrange(AIC)
```


Based on the AIC, model m5.1 looks to be the best model for our data.

## Inspect model m5.1

```{r summarise-m5}
appraise(m5.1)
summary(m5.1)
```

```{r visualise-gam}
draw(m5.1)
```

## Predict spore dispersal

```{r predict}
mod_dat %>% distinct(downwind) %>% pull()
newd <- expand_grid(
  time_min = c(90, 180, 270),
  downwind = c("downwind", "lateral", "upwind"),
  distance_m = seq(0, 400, 100),
  wind_speed = 0:30,
  field = 5
)
newd$pred_spore_cm2 <- predict.gam(m5.1, newd)
```

## Visualise spore dispersal predictions

```{r graph-predictions, fig.cap="Predictions of spore dispersal at 0 to 400m. Points indicate median values with upper and lower 25% quartile and 75% quartiles represented with vertical lines with a smoothed line fit through the points."}
newd %>%
  ggplot(aes(x = distance_m, y = pred_spore_cm2, col = downwind)) +
  stat_summary(
    mapping = aes(x = distance_m, y = pred_spore_cm2),
    fun.min = function(z) {
      quantile(z, 0.25)
    },
    fun.max = function(z) {
      quantile(z, 0.75)
    },
    fun = median
  ) +
geom_smooth() +
  scale_fill_viridis_d("Prevailing wind") +
  scale_colour_viridis_d("Prevailing Wind") +
  facet_grid(. ~ time_min) +
  scale_x_continuous(sec.axis = sec_axis(
    ~ . ,
    name = "Minutes since harvest beginning",
    breaks = NULL,
    labels = NULL
  )) +
  labs(x = "Distance from harvesting field (m)",
       y = "Spores / m²") +
  theme(
    legend.position = "top",
    legend.justification = "center",
    legend.direction = "horizontal",
    legend.background = element_blank(),
    legend.key = element_blank()
  ) +
  guides(fill = guide_legend(title.position = "top",
                             title.hjust = 0.5))
```

```{r save-prediction-graph, eval=FALSE}
ggsave(
  last_plot(),
  file = "plots/spore_prediction.png",
  w = 5,
  h = 2,
  scale = 1.2
)
```



