---
title: "Fit models"
output:
  html_document:
    theme: simplex
    toc: true
    toc_float:
      collapsed: no
      smooth_scroll: yes
    code_folding: hide
    toc_depth: 2
    css: style.css
---

```{r markdown-settings, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 7,
  message = FALSE,
  warning = FALSE
)
```

## Load libraries

```{r session settings}
library("tidyverse")
library("broom")
library("ggplot2")
library("ggpubr")
library("gratia")
library("here")
library("mgcv")

theme_set(theme_pubclean())
```

## Load data

```{r load-data}
load(here("data/mod_dat.Rdata"))
```

## Inspect the mod_dat object

```{r inspect-mod-dat}
mod_dat %>%
  distinct(field, trap_coord, .keep_all = TRUE) %>%
  select(trap_coord,
         trap_degrees,
         wind_degrees,
         degree_dif,
         downwind,
         trap_id) %>%
  print(n = Inf)
```

Because of the number of zeros in the data, we will use a [Tweedie family](https://en.wikipedia.org/wiki/Tweedie_distribution) for the models to help account for this, _i.e._ in the model functions you will see `family = "tw"` for this.

### Model m1

* Response: `spore_cm2`
* Linear predictors: `time_slice`, `distance_m`, `downwind`
* Smoothed predictors: `wind_speed`
* Random effects: `trap_id`

```{r m1}
m1 <- gam(
  spore_cm2 ~ time_slice + distance_m + downwind +
    s(wind_speed, k = 18) +
    s(trap_id, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m1
```

### Model m2

* Response: `spore_cm2`
* Linear predictors: `time_slice`, `distance_m`, `downwind`
* Smoothed predictors: `wind_speed`
* Random effects: `trap_id`, `field`

```{r m2}
m2 <- gam(
  spore_cm2 ~ time_slice + distance_m + downwind +
    s(wind_speed, k = 18) +
    s(trap_id, bs = "re") +
    s(field, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m2
```

### Model m3

* Response: `spore_cm2`
* Linear predictors: `time_slice`, `distance_m`, `downwind`
* Smoothed predictors: `wind_speed`
* Random effects: `trap_id`, `time_slice`

```{r m3}
m3 <- gam(
  spore_cm2 ~ time_slice + distance_m + downwind +
    s(wind_speed, k = 18) +
    s(trap_id, bs = "re") +
    s(time_slice, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m3
```

### Model m4

* Response: `spore_cm2`
* Linear predictors: `distance_m`, `downwind`
* Smoothed predictors: `wind_speed`
* Random effects: `trap_id`, `time_slice`

```{r m4}
m4 <- gam(
  spore_cm2 ~ distance_m + downwind +
    s(wind_speed, k = 18) +
    s(trap_id, bs = "re") +
    s(time_slice, bs = "re"),
  method = "REML",
  data = mod_dat,
  family = "tw"
)
m4
```

## Model AIC

Check models for best fit using AIC.

```{r AIC}
models <- list(
  m1 = m1,
  m2 = m2,
  m3 = m3,
  m4 = m4
)
map_df(models, glance, .id = "model") %>%
  arrange(AIC, BIC)
```

Based on the AIC and BIC values, model m4 looks to be the best model for our data.

## Inspect Model m4

### Check model

```{r gam-check-m4}
gam.check(m4)
```

The model fully converged after 9 iterations.
All of the predictive factors included were significant at _P_ > 0.05 and the smoothed predictors indicate a good fit with residuals being randomly distributed and enough basis functions for them to be properly fit.
The Q-Q plot generally falls along the line indicating a sufficient fit.
There is a bit of a pattern in the residual vs linear predictors but the histogram of residuals is a nice bell-shape.
The response versus fitted values clusters nicely around the 1:1 line.
In all, this model looks to explain the spore dispersal data very well.

### Summarise and visualise model m4

```{r appraise-m4}
summary(m4)
plot(m4)
```

## Predict spore dispersal

```{r predict}
newd <- expand_grid(
  time_slice = c(90, 180, 270),
  downwind = c("downwind", "lateral", "upwind"),
  distance_m = seq(0, 400, 100),
  wind_speed = 0:30,
  field = 1,
  trap_id = as.factor(apply(expand.grid(
    1, c("_N_", "_S_", "_E_", "_W_"), c(100, 200, 300, 400)
  ),
  1, paste, collapse = "")),
  sub_sample = as.factor(c(1, 2))
)
  
newd$pred_spore_cm2 <- predict(m4, newd)
```

## Visualise spore dispersal predictions

Visualise the spore dispersal predictions for 0m, 100m, 200m, 300m, 400m for traps placed upwind, lateral and downwind of the harvest field for sampling times at 90, 180 and 270 minutes after harvest started.

```{r graph-predictions}
newd %>%
  mutate(pred_spore_cm2 = replace(pred_spore_cm2, pred_spore_cm2 < 0, 0)) %>%
  ggplot(aes(x = as.factor(distance_m), y = pred_spore_cm2)) +
  geom_violin() +
  facet_grid(. ~ downwind) +
  labs(x = "Distance from harvesting field (m)",
       y = "Predicted Spores / m²"
  )
```

```{r save-prediction-graph, eval=FALSE}
ggsave(
  last_plot(),
  file = "plots/spore_prediction.png",
  w = 5,
  h = 2,
  scale = 1.2
)
```
